---
title: "Homework 2"
author: "Yueran Pan"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)
library(tidyverse)
```

## Problem 1
```{r import_subway}
subway_df = read_csv(
    "hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
head(subway_df)
```


```{r read_clean}

names(subway_df)
#clean names to lower case
subway_df = janitor::clean_names(subway_df)
#check new name
names(subway_df)

#Keep the column specified
subway_df = select (
    subway_df, 
    line:entry, 
    vending,
    ada
    )

#Convert the entry variable from character to logical variable using case_match function

subway_df = mutate(subway_df,
  entry = case_match(
    entry, 
    "YES" ~ TRUE,
    "NO"~ FALSE
  )
)
```


About this dataset:

The dataset originally contained records of NYC subway entrances and exits, including line, station name, routes served, entry permission, vending availability, and so on. I cleaned the column names to lower case, retained the specified columns (line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance), and converted column `entry` from a character variable to a logical variable. The resulting dataset has 1,868 rows and 19 columns. It is not yet completely tidy because route names are spread across multiple columns, from route1 to route11.

###Prob1,Q1:How many distinct stations are there?
```{r count_stations}
stations_df = distinct(
  subway_df,
  line,
  station_name
)
str(stations_df)

```

There are 464 distinct stations, identified by the combination of line and station name.


## Problem 2

## Problem 3