---
title: "Homework 2"
author: "Yueran Pan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)
library(tidyverse)
```

## Problem 1
```{r import_subway}
 subway_df = 
  read_csv(file = "./hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  na = c(".", "NA", ""),
  col_types = cols(
     Route8 = col_character(),
     Route9 = col_character(),
     Route10 = col_character(),
     Route11 = col_character()
     )
 )
```


```{r read_clean}

names(subway_df)
#clean names to lower case
subway_df = janitor::clean_names(subway_df)
#check new name
names(subway_df)

#Keep the column specified
subway_df = select (
    subway_df, 
    line:entry, 
    vending,
    ada
    )

#Convert the entry variable from character to logical variable using case_match function

subway_df = mutate(subway_df,
  entry = case_match(
    entry, 
    "YES" ~ TRUE,
    "NO"~ FALSE
  )
)
```


About this dataset:

The dataset originally contained records of NYC subway entrances and exits,  including 32 variables, such as line, station name, routes served, entry permission, vending availability, and so on. I cleaned the column names to lower case, retained the specified columns (line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance), and converted column `entry` from a character variable to a logical variable. The resulting dataset has 1,868 rows and 19 columns. It is not yet completely tidy because route names are spread across multiple columns, from route1 to route11.

### Prob1, Q1: How many distinct stations are there?
```{r count_stations}
stations_df = distinct(
  subway_df,
  line,
  station_name
)
str(stations_df)

```
There are 465 distinct stations, identified by the combination of line and station name.



### Prob1, Q2: How many stations are ADA compliant?
```{r count_ada_stations}
ada_df = filter(subway_df, ada == TRUE)

ada_df = distinct(
  ada_df,
  ada,
  line,
  station_name)
str(ada_df)
```
There are 84 ADA compliant stations.


### Prob1,Q3: What proportion of station entrances / exits without vending allow entrance?
```{r entry_and_count_without_vending }
wo_vending_df = filter(
  subway_df,
  vending == "NO"
)
wo_vend_entry = pull(wo_vending_df,entry)
mean(wo_vend_entry == TRUE)
```
37.70% of station entrances/exits without vending allow entrance.


### Prob1,Q4: Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r a_train_stations}
# Reshape route columns into route number and route name.
subway_long_df = pivot_longer(
  subway_df,
  route1:route11,
  names_to = "route_number",
  names_prefix = "route",
  values_to = "route_name"
)
subway_long_df = drop_na(subway_long_df, route_name)

# Keep records serving the A train.
a_train_df = filter(
  subway_long_df,
  route_name == "A"
)
# Count distinct A train stations.
a_stations_df = distinct(
  a_train_df,
  line,
  station_name
)
#check
str(a_stations_df)

# Count distinct ADA-compliant A train stations.
a_ada_df = filter(
  a_train_df,
  ada == TRUE
)
a_ada_stations_df = distinct(
  a_ada_df,
  line,
  station_name
)
#final check
str(a_ada_stations_df)
```
Among all 60 A train stations served, 17 of them are ADA compliant.



## Problem 2
```{r import_trash_wheels}

mr_df = readxl::read_excel(
 "./hw2_data/202610 Trash Wheel Collection Data.xlsx",
 sheet = "Mr. Trash Wheel",
range = "A2:N742"
)|>
janitor::clean_names()|>
drop_na(dumpster) |>
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls),
  year = as.integer(year),
  plastic_bottles = as.integer(plastic_bottles),
  wheel= "Mr. Trash Wheel"
  )

prof_df = readxl::read_excel(
 "./hw2_data/202610 Trash Wheel Collection Data.xlsx",
 sheet = "Professor Trash Wheel",
range = "A2:M139",
#saw a "dive" in G137 of prof sheet...???Thus I treat this nonnumeric value "dive" as missing. 
na = c("", "dive")
)|>
janitor::clean_names()|>
drop_na(dumpster) |>
mutate(
  year = as.integer(year),
  plastic_bottles = as.integer(plastic_bottles),
  wheel= "Professor Trash Wheel")

gwynnda_df = readxl::read_excel(
 "./hw2_data/202610 Trash Wheel Collection Data.xlsx",
 sheet = "Gwynnda the Good Wheel of the W",
range = "A2:L394",
)|>
janitor::clean_names()|>
drop_na(dumpster) |>
mutate(
  year = as.integer(year),
  plastic_bottles = as.integer(plastic_bottles),
  wheel= "Gwynnda")
```

### combine 3 sheets

```{r combine_3_wheels} 
trash_df = bind_rows(
  mr_df,
  prof_df,
  gwynnda_df
)
str(trash_df)
```

### count 
```{r total_trash} 
prof_wt_just_in_case = drop_na(prof_df, weight_tons)
prof_wt = pull(prof_wt_just_in_case, weight_tons)
prof_sum_wt = mean(prof_wt)*length(prof_wt)
prof_sum_wt

gwynnda_June22 = filter(
   gwynnda_df,
   month == "June",
   year == 2022
)

gwynnda_June22 = drop_na(gwynnda_June22, cigarette_butts)
gwynnda_cigbt = pull(gwynnda_June22, cigarette_butts)

gwynnda_count_cigbt = mean(gwynnda_cigbt)*length(gwynnda_cigbt)
gwynnda_count_cigbt
```

The total weight of trash collected by Professor Trash Wheel was 294.96 tons.
18120 cigarette butts were collected by Gwynnda in June of 2022.



## Problem 3

### Import&read baseline, excluding those with MCI already
```{r mci_baseline}
baseline_raw_df = read_csv(
  "./hw2_data/MCI_baseline.csv",
  skip = 1,
  na = c(".", "NA", "")
) |>
  janitor::clean_names()

length(pull(baseline_raw_df, id))

baseline_df = baseline_raw_df |>
  filter(
    is.na(age_at_onset) | age_at_onset > current_age
  ) |>
  mutate(
    sex = case_match(
      sex,
      0 ~ "Female",
      1 ~ "Male"
    ),
    apoe4 = case_match(
      apoe4,
      0 ~ "Non-carrier",
      1 ~ "Carrier"
    )
  )

pull(baseline_df, id)|>
  length()

mci_cases_df = baseline_df |>
  filter(!is.na(age_at_onset))

pull(mci_cases_df, id)|>
  length()

pull(baseline_df, current_age)|>
  mean()

women_df = baseline_df |>
  filter(sex == "Female")

mean(pull(women_df, apoe4) == "Carrier")
```
This part, I skipped the first explanatory row and cleaned out missing values (periods). Then I formatted sex and APOE4 carrier status by character (for sex where 0=female, 1=sex, for apoe4 where 0=non-carrier, 1=carrier). 
Among the 483 recruited participants, 4 with MCI onset before the baseline age were excluded. 
93 developed MCI during follow-up among the remaining 479 eligible participants, with the baseline age 65.03 year-old, with 30% of women being APOE4 carriers.

```{r}
amyloid_df = read_csv(
  "./hw2_data/mci_amyloid.csv",
  skip = 1,
  na = c(".", "NA", "")
) |>
  janitor::clean_names() |>
  rename(id = study_id)
#in order to merge two sheets, keep name consistent. chose 'id' from bsl, renaming amyld's.

amyloid_long_df = amyloid_df |>
  pivot_longer(
    baseline:time_8,
    names_to = "visit",
    values_to = "amyloid_ratio"
  ) |>
  mutate(
    visit = case_match(
      visit, 
      "baseline" ~ 0,
      "time_2" ~ 2,
      "time_4" ~ 4,
      "time_6" ~ 6,
      "time_8" ~ 8
    )
  )

str(amyloid_long_df)
```
Continue on, like how I treated baseline.csv, I convert the column title variables to lower cases and renamed participant identified to id to keep variables consistent. I transposed the five measurement columns into visit and amyloid ratio, then coded subsequent visit year since baseline. The resulting dataset contains 2435 partipants' visit records from 487 participants. And because I `pivot_long`, any missing  biomarker measurements were remained, shown as NA (Each participant has 5 visit slots)

```{r crosscheck_ids}
baseline_ids = pull(baseline_df, id)
amyloid_ids = pull(amyloid_df, id)

baseline_only_df = baseline_df |>
  filter(!(id %in% amyloid_ids))

amyloid_only_df = amyloid_df |>
  filter(!(id %in% baseline_ids))

length(pull(baseline_only_df, id))
length(pull(amyloid_only_df, id))
```
8 participants appeared only in the baseline datasets. Another 16 participants appeared only in the amyloid datasets.

```{r merge_those_appear_in_both_amy_and_base}

baseline_matched_df = baseline_df |>
  filter(id %in% amyloid_ids)

merge_mci_df = left_join(
  baseline_matched_df,
  amyloid_long_df,
  by = "id"
)       

length(pull(baseline_matched_df, id))
str(merge_mci_df)

#export csv
write_csv(merge_mci_df,"./hw2_data/mci_merged.csv")
```

As a result, I obtained 471 partcipants from the merged data, combining their baseline characteristics with the longitudinal biomarker records by id. There were 2,355 participants' visit records, and 8 variables [2.355 x 8]. Csv were exported.