---
title: "Homework 2"
author: "Yueran Pan"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)
library(tidyverse)
```

## Problem 1
```{r import}
 subway_df = 
  read_csv(file = "./hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  na = c(".", "NA", ""),
  col_types = cols(
     Route8 = col_character(),
     Route9 = col_character(),
     Route10 = col_character(),
     Route11 = col_character()
     )
 )
```


```{r read_clean}

names(subway_df)
#clean names to lower case
subway_df = janitor::clean_names(subway_df)
#check new name
names(subway_df)

#Keep the column specified
subway_df = select (
    subway_df, 
    line:entry, 
    vending,
    ada
    )

#Convert the entry variable from character to logical variable using case_match function

subway_df = mutate(subway_df,
  entry = case_match(
    entry, 
    "YES" ~ TRUE,
    "NO"~ FALSE
  )
)
```


About this dataset:

The dataset originally contained records of NYC subway entrances and exits,  including 32 variables, such as line, station name, routes served, entry permission, vending availability, and so on. I cleaned the column names to lower case, retained the specified columns (line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance), and converted column `entry` from a character variable to a logical variable. The resulting dataset has 1,868 rows and 19 columns. It is not yet completely tidy because route names are spread across multiple columns, from route1 to route11.

### Prob1, Q1: How many distinct stations are there?
```{r count_stations}
stations_df = distinct(
  subway_df,
  line,
  station_name
)
str(stations_df)

```
There are 465 distinct stations, identified by the combination of line and station name.



### Prob1, Q2: How many stations are ADA compliant?
```{r count_ada_stations}
ada_df = filter(subway_df, ada == TRUE)

ada_df = distinct(
  ada_df,
  ada,
  line,
  station_name)
str(ada_df)
```
There are 84 ADA compliant stations.


### Prob1,Q3: What proportion of station entrances / exits without vending allow entrance?
```{r entry_and_count_without_vending }
wo_vending_df = filter(
  subway_df,
  vending == "NO"
)
wo_vend_entry = pull(wo_vending_df,entry)
mean(wo_vend_entry == TRUE)
```
37.70% of station entrances/exits without vending allow entrance.


### Prob1,Q4: Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r a_train_stations}
# Reshape route columns into route number and route name.
subway_long_df = pivot_longer(
  subway_df,
  route1:route11,
  names_to = "route_number",
  names_prefix = "route",
  values_to = "route_name"
)
subway_long_df = drop_na(subway_long_df, route_name)

# Keep records serving the A train.
a_train_df = filter(
  subway_long_df,
  route_name == "A"
)
# Count distinct A train stations.
a_stations_df = distinct(
  a_train_df,
  line,
  station_name
)
#check
str(a_stations_df)

# Count distinct ADA-compliant A train stations.
a_ada_df = filter(
  a_train_df,
  ada == TRUE
)
a_ada_stations_df = distinct(
  a_ada_df,
  line,
  station_name
)
#final check
str(a_ada_stations_df)
```
Among all 60 A train stations served, 17 of them are ADA compliant.



## Problem 2

## Problem 3